---
title: "Initial Korean covid 19"
author: "Kaylie Fagoaga"
date: "Summer 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

# Introduction

Is there a correlation between death and gender? 

### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female  
77 of the data entries were coded sex  as blank, so we omit these.  
state of a patient must be died or alive - we adjust that as well.    
  
# Your Question - Is dying independent of sex?

$H_0$: There is no difference in the death rates depending on sex.

$H_A$: There is a difference in the death rates depending on sex.
 
# Methods 

result is a categorical variable, sex is categorical   
result depends on sex, we use helper file
cat~cat 

# Results 


## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

This barchart shows that almost twice as much males died than females 





  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This numerical summary shows that out of 100% of people who died, 64% male died and 36% female died which indicates that the death of males almost doubled than death in females.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi-squared test is 0.001, so we reject the null hypothesis that gender and death by Covid are independent, they are dependent.   
The Fisher exact test shows that men are 2.3 times more likely to die than women in Korea. 

## Conclusions
The original question was is there a correlation between death and gender?        
The results were that there is a significance relationship between gender and death.   
P-value was 0.001 which meant we had to reject the null hypothesis because it was less than 5.        
The death ratio from men to women was 2.3 to 1 which proved more deaths in men than women.  
Men in Korea are most likely to die from Covid than women in Korea. This means sex and death are dependent. 
 